# Supervised Learning In R Regression
***
Notes taken during/inspired by the DataCamp course 'Supervised Learning In R Classification' by Nina Zumel and John Mount.

**_Course Handouts_**

* [Part 1 - What is Regression?](./files/SLinRRegression/chapter1.pdf)


**_Other useful links_**

* []()

## What is Regression?

In this course we are interested in predicting numerical outcomes (rather than discreet) from a set of input variables.  Classification on the other hand is the task of making discreet predictions.  WE can break down the modelling process in to two camps:

* Scientific mindset - we try to understand causal mechanisms 
* Engineering mindset - we try to accuratley predict

Machine Learning is more in line with the Engineering mindset.

In the first task, the goal is to predict the rate of female unemployment from the observed rate of male unemployment. The outcome is female_unemployment, and the input is male_unemployment.

```{r}
# Loand the unemployment data
unemployment <- readRDS("./files/SLinRRegression/unemployment.rds")

# unemployment is loaded in the workspace
summary(unemployment)

# Define a formula to express female_unemployment as a function of male_unemployment
fmla <- as.formula("female_unemployment ~ male_unemployment")

# Print it
fmla

# Use the formula to fit a model: unemployment_model
unemployment_model <- lm(formula = fmla, data = unemployment)

# Print it
unemployment_model
```

We can see the relationship is positive - female unemployment increases as male unemployment does.  We can then use some diagnostics on the model.  There are different ways and packages to achieve this.

```{r}
# broom and sigr are already loaded in your workspace
# Print unemployment_model
unemployment_model

# Call summary() on unemployment_model to get more details
summary(unemployment_model)

# Call glance() on unemployment_model to see the details in a tidier form
broom::glance(unemployment_model)

# Call wrapFTest() on unemployment_model to see the most relevant details
sigr::wrapFTest(unemployment_model)
```

In the next section we use the unemployment model unemployment_model to make predictions from the unemployment data, and compare predicted female unemployment rates to the actual observed female unemployment rates on the training data, unemployment. You will also use your model to predict on the new data in newrates.

```{r}
newrates <- data.frame(male_unemployment = 5)

# Predict female unemployment in the unemployment data set
unemployment$prediction <-  predict(unemployment_model, unemployment)

# load the ggplot2 package
library(ggplot2)

# Make a plot to compare predictions to actual (prediction on x axis). 
ggplot(unemployment, aes(x = prediction, y = female_unemployment)) + 
  geom_point() +
  geom_abline(color = "blue")

# Predict female unemployment rate when male unemployment is 5%
pred <- predict(unemployment_model, newrates)
# Print it
pred
```

Next we will look at a model of blood pressure as a function of age and weight.

```{r}
# Load the blood data
bloodpressure <- readRDS("./files/SLinRRegression/bloodpressure.rds")
```

